---
title: "Lab 4a: Data Visualization with ggplot2"
author: "Abigail Gray"
date: "September 20, 2020"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library("tidyverse")
```


```{r}
surveys_complete <- read_csv("data/surveys_complete.csv")
```


### Plotting with ggplot2

```{r}
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) + geom_point()
```
```{r}
surveys_plot <- ggplot(data = surveys_complete, 
                       mapping = aes(x = weight, y = hindfoot_length))

surveys_plot + 
    geom_point()
```


```{r}
#install.packages("hexbin")
library("hexbin")
```
### Challenge (optional)
```{r}
surveys_plot +
  geom_hex()
```
* Strengths: shows density of points in different areas of graph when there is overplotting of points, weaknesses: you cannot see all the spaces between points in areas on graph where there is not a lot of overlap


### Building your plots iteratively
```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1)
```
```{r}
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1, color = "blue")
```
```{r}
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1, aes(color = species_id))
```

### Challenge

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_point(aes(color = plot_type))
```
* This isn't a good way to show the data, scatter plots are meant to show continuous data but species_id is a concrete variable, so the data is broken up into separate columns and its difficult to interpret the data. It would be better to use a boxplot.


### Boxplot

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_boxplot()
```

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
  geom_boxplot(alpha = 0) +
  geom_jitter(alpha = 0.3, color = "tomato")
```
* To put boxplot in front of points so that it's not hidden, put geom_jitter() before geom_boxplot()

### Challenge

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_violin()
```
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_violin() +
    scale_y_log10()
```

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(aes(color = plot_id)) +
  geom_boxplot()
```
```{r}
new_plot_id <- factor(surveys_complete$plot_id)

ggplot(data = surveys_complete, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(aes(color = new_plot_id)) +
  geom_boxplot()
```


* When plot_id is an interger, the color of the data points are on a gradient scale where 1 is the lightest shade and 24 is the darkest shade. When plot_id is a factor, each plot_id number is assigned a different color because they are treated as separate levels.


### Plotting time series data

```{r}
yearly_counts <- surveys_complete %>% 
  count(year, genus)
```


```{r}
ggplot(data = yearly_counts, aes(x = year, y = n)) +
  geom_line()
```


```{r}
ggplot(data = yearly_counts, aes(x = year, y = n, group = genus)) +
  geom_line()
```


```{r}
ggplot(data = yearly_counts, aes(x = year, y = n, color = genus)) +
  geom_line()
```


### Integrating the pipe operator with ggplot2

```{r}
yearly_counts %>% 
    ggplot(mapping = aes(x = year, y = n, color = genus)) +
    geom_line()
```


```{r}
yearly_counts_graph <- surveys_complete %>%
    count(year, genus) %>% 
    ggplot(mapping = aes(x = year, y = n, color = genus)) +
    geom_line()

yearly_counts_graph
```


### Faceting

```{r}
ggplot(data = yearly_counts, aes(x = year, y = n)) +
  geom_line() +
  facet_wrap(facets = vars(genus))
```


```{r}
yearly_sex_counts <- surveys_complete %>%
  count(year, genus, sex)
```


```{r}
ggplot(data = yearly_sex_counts, mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_wrap(facets =  vars(genus))
```


```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(rows = vars(sex), cols =  vars(genus))
```


```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(rows = vars(genus))
```


```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(cols = vars(genus))
```


### ggplot2 themes

```{r}
 ggplot(data = yearly_sex_counts, 
        mapping = aes(x = year, y = n, color = sex)) +
     geom_line() +
     facet_wrap(vars(genus)) +
     theme_bw()
```


### Challenge

```{r}
yearly_weight <- surveys_complete %>%
  group_by(year, species_id) %>% 
  summarize(average_weight = mean(weight))
```
```{r}
ggplot(data = yearly_weight, mapping = aes(x = year, y = average_weight)) +
  geom_line() +
  facet_wrap(vars(species_id)) +
  theme_bw()
```


### Customization

```{r}
ggplot(data = yearly_sex_counts, aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus)) +
    labs(title = "Observed genera through time",
         x = "Year of observation",
         y = "Number of individuals") +
    theme_bw()
```


```{r}
ggplot(data = yearly_sex_counts, mapping = aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus)) +
    labs(title = "Observed genera through time",
        x = "Year of observation",
        y = "Number of individuals") +
    theme_bw() +
    theme(text = element_text(size = 16))
```
```{r}
#install.packages('extrafont')
```
```{r}
library(extrafont)
```


```{r}
ggplot(data = yearly_sex_counts, mapping = aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus)) +
    labs(title = "Observed genera through time",
        x = "Year of observation",
        y = "Number of individuals") +
    theme_bw() +
    theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 12),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 16))
```


```{r}
grey_theme <- theme(axis.text.x = element_text(colour="grey20", size = 12, 
                                               angle = 90, hjust = 0.5, 
                                               vjust = 0.5),
                    axis.text.y = element_text(colour = "grey20", size = 12),
                    text=element_text(size = 16))

ggplot(surveys_complete, aes(x = species_id, y = hindfoot_length)) +
    geom_boxplot() +
    grey_theme
```


### Challenge

```{r}
ggplot(data = yearly_sex_counts, mapping = aes(x = year, y = n, color = sex)) +
    geom_line(size = 2) +
    facet_wrap(vars(genus)) +
    labs(title = "Observed genera through time",
        x = "Year of observation",
        y = "Number of individuals") +
    theme_bw() +
    theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 12),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 16)) +
  scale_color_discrete(name = "Sex",
                       labels = c("Female", "Male"))
  
```


### Arranging and exporting plots


```{r}
library(gridExtra)
```

```{r}
spp_weight_boxplot <- ggplot(data = surveys_complete, 
                             aes(x = species_id, y = weight)) +
  geom_boxplot() +
  labs(x = "Species", 
       y = expression(log[10](Weight))) +
  scale_y_log10() + 
  labs()

spp_count_plot <- ggplot(data = yearly_counts, 
                         aes(x = year, y = n, color = genus)) +
  geom_line() + 
  labs(x = "Year", y = "Abundance")

grid.arrange(spp_weight_boxplot, spp_count_plot, ncol = 2, widths = c(4, 6))
```

```{r}
my_plot <- ggplot(data = yearly_sex_counts, 
                  aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus)) +
    labs(title = "Observed genera through time",
        x = "Year of observation",
        y = "Number of individuals") +
    theme_bw() +
    theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90,
                                     hjust = 0.5, vjust = 0.5),
          axis.text.y = element_text(colour = "grey20", size = 12),
          text = element_text(size = 16))

ggsave("lab4a_abigailgray.png", my_plot, width = 15, height = 10)
```


```{r}
combo_plot <- grid.arrange(spp_weight_boxplot, spp_count_plot, ncol = 2, 
                           widths = c(4, 6))
ggsave("combo_plot_abun_weight.png", combo_plot, width = 10, dpi = 300)
```










