---
title: "Lab 4: Data wrangling and graphing COVID-19 reporting data (Part 1)"
author: "Abigail Gray"
date: "September 20, 2020"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r}
report_03_11_2020 <- read_csv(url("https://github.com/CSSEGISandData/COVID-19/raw/master/csse_covid_19_data/csse_covid_19_daily_reports/03-11-2020.csv")) %>% rename(Country_Region = "Country/Region", Province_State = "Province/State")
```
```{r}
str(report_03_11_2020)
```
```{r}
view(report_03_11_2020)
```

```{r}
#install.packages('DT')
```

```{r}
xfun::session_info('DT')
```

```{r}
library(DT)
datatable(report_03_11_2020)
```

### Exercises


### Question 1

```{r}
report_03_12_2020 <- read_csv(url("https://github.com/CSSEGISandData/COVID-19/raw/master/csse_covid_19_data/csse_covid_19_daily_reports/03-12-2020.csv")) %>% rename(Country_Region = "Country/Region", Province_State = "Province/State")
```
```{r}
str(report_03_12_2020)
view(report_03_12_2020)
```
```{r}
report_09_12_2020 <- read_csv(url("https://github.com/CSSEGISandData/COVID-19/raw/master/csse_covid_19_data/csse_covid_19_daily_reports/09-12-2020.csv"))
```
```{r}
str(report_09_12_2020)
view(report_09_12_2020)
```

* Six more columns were added between 3/12/2020 and 9/12/2020 including FIPS, Admin2, Active, Combined_Key, Incidence_Rate, and Case-Fatality_Ratio.



### Question 2

```{r}
report_03_12_2020 %>%
  group_by(Country_Region) %>% 
  summarize(count = n())
```

```{r}
report_09_12_2020 %>%
  group_by(Country_Region) %>% 
  summarize(count = n())
```



### Question 3
```{r}
report_03_12_2020 %>% 
  filter(Country_Region == 'US') %>% 
  group_by(Province_State) %>% 
  summarize(count = n())
```

```{r}
report_09_12_2020 %>% 
  filter(Country_Region == 'US') %>% 
  group_by(Province_State) %>% 
  summarize(count = n())
```



### Question 4

```{r}
TopDeaths_03_12 <- report_03_12_2020 %>%
  group_by(Country_Region) %>% 
  summarize(Deaths_by_Country = sum(Deaths)) %>% 
  arrange(desc(Deaths_by_Country)) %>% 
  slice(1:10)
TopDeaths_03_12
```

```{r}
TopDeaths_09_12 <- report_09_12_2020 %>%
  group_by(Country_Region) %>% 
  summarize(Deaths_by_Country = sum(Deaths)) %>% 
  arrange(desc(Deaths_by_Country)) %>% 
  slice(1:10)
TopDeaths_09_12
```



### Question 5
```{r}
TopConfirmed_03_12 <- report_03_12_2020 %>%
  group_by(Country_Region) %>% 
  summarize(Confirmed_by_Country = sum(Confirmed)) %>% 
  arrange(desc(Confirmed_by_Country)) %>% 
  slice(1:10)
TopConfirmed_03_12
```
```{r}
TopConfirmed_09_12 <- report_09_12_2020 %>%
  group_by(Country_Region) %>% 
  summarize(Confirmed_by_Country = sum(Confirmed)) %>% 
  arrange(desc(Confirmed_by_Country)) %>% 
  slice(1:10)
TopConfirmed_09_12
```
* Ten Countries with Most Confirmed Cases on 03/12/20
```{r}
ggplot(data = TopConfirmed_03_12, aes(x = Country_Region, y = Confirmed_by_Country)) +
  geom_point()
```



* Ten Countries with Most Deaths on 03/12/2020 
```{r}
ggplot(data = TopDeaths_03_12, aes(x = Country_Region, y = Deaths_by_Country)) +
  geom_point()
```



* Ten Countries with Most Confirmed Cases on 09/12/20
```{r}
ggplot(data = TopConfirmed_09_12, aes(x = Country_Region, y = Confirmed_by_Country)) +
  geom_point()
```



* Ten Countries with Most Deaths on 09/12/2020
```{r}
ggplot(data = TopDeaths_09_12, aes(x = Country_Region, y = Deaths_by_Country)) +
  geom_point()
```



### Question 6

```{r}
TopStateConfirmed_03_12 <- report_03_12_2020 %>% 
  filter(Country_Region == 'US') %>% 
  group_by(Province_State) %>%
  summarize(Confirmed_by_State = sum(Confirmed)) %>% 
  arrange(desc(Confirmed_by_State)) %>% 
  slice(1:10)
TopStateConfirmed_03_12
```

Ten U.S. States with Most Confirmed Cases on 03/12/2020

```{r}
ggplot(data = TopStateConfirmed_03_12, aes(x = Confirmed_by_State, y = Province_State)) +
  geom_bar(stat = "identity")
```


```{r}
TopStateDeaths_03_12 <- report_03_12_2020 %>% 
  filter(Country_Region == 'US') %>% 
  group_by(Province_State) %>%
  summarize(Deaths_by_State = sum(Deaths)) %>% 
  arrange(desc(Deaths_by_State)) %>% 
  slice(1:10)
TopStateDeaths_03_12
```

* Ten U.S. States with Most Deaths on 03/12/2020

```{r}
ggplot(data = TopStateDeaths_03_12, aes(x = Deaths_by_State, y = Province_State)) +
  geom_bar(stat = "identity")
```



```{r}
TopStateConfirmed_09_12 <- report_09_12_2020 %>% 
  filter(Country_Region == 'US') %>% 
  group_by(Province_State) %>%
  summarize(Confirmed_by_State = sum(Confirmed)) %>% 
  arrange(desc(Confirmed_by_State)) %>% 
  slice(1:10)
TopStateConfirmed_09_12
```

* Top U.S. States with Confirmed Cases on 09/12/2020

```{r}
ggplot(data = TopStateConfirmed_09_12, aes(x = Confirmed_by_State, y = Province_State)) +
  geom_bar(stat = "identity")
```


```{r}
TopStateDeaths_09_12 <- report_09_12_2020 %>% 
  filter(Country_Region == 'US') %>% 
  group_by(Province_State) %>%
  summarize(Deaths_by_State = sum(Deaths)) %>% 
  arrange(desc(Deaths_by_State)) %>% 
  slice(1:10)
TopStateDeaths_09_12
```

* Ten U.S States with Most Deaths on 09/12/2020

```{r}
ggplot(data = TopStateDeaths_09_12, aes(x = Deaths_by_State, y = Province_State)) +
  geom_bar(stat = "identity")
```

