---
title: "Lab 4a Data Visualization with ggplot2"
author: "Abigail Gray"
date: "September 20, 2020"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library("tidyverse")
```


```{r}
surveys_complete <- read_csv("data/surveys_complete.csv")
```

```{r}
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) + geom_point()
```
```{r}
surveys_plot <- ggplot(data = surveys_complete, 
                       mapping = aes(x = weight, y = hindfoot_length))

surveys_plot + 
    geom_point()
```


```{r}
#install.packages("hexbin")
library("hexbin")
```
### Challenge (optional)
```{r}
surveys_plot +
  geom_hex()
```
* Strengths: shows density of points in different areas of graph when there is overplotting of points, weaknesses: you cannot see all the spaces between points in areas on graph where there is not a lot of overlap

```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1)
```
```{r}
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1, color = "blue")
```
```{r}
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1, aes(color = species_id))
```

### Challenge

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_point(aes(color = plot_type))
```
* This isn't a good way to show the data, scatter plots are meant to show continuous data but species_id is a concrete variable, so the data is broken up into separate columns and its difficult to interpret the data. It would be better to use a boxplot.

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_boxplot()
```

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
  geom_boxplot(alpha = 0) +
  geom_jitter(alpha = 0.3, color = "tomato")
```
* To put boxplot in front of points so that it's not hidden, put geom_jitter() before geom_boxplot()

### Challenge

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_violin()
```
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_violin() +
    scale_y_log10()
```

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(aes(color = plot_id)) +
  geom_boxplot()
```
```{r}
new_plot_id <- factor(surveys_complete$plot_id)

ggplot(data = surveys_complete, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(aes(color = new_plot_id)) +
  geom_boxplot()
```


* When plot_id is an interger, the color of the data points are on a gradient scale where 1 is the lightest shade and 24 is the darkest shade. When plot_id is a factor, each plot_id number is assigned a different color because they are treated as separate levels.




