---
title: 'Lab 9A: Programming in R - Regular Expressions'
author: "Abigail Gray"
date: "10/29/2020"
output: html_document
---

## 14.3 Matching patterns with regular expressions

```{r message=FALSE, warning=FALSE}
library(stringr)
library(tidyverse)
```

### 14.3.1 Basic matches

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
```

```{r}
str_view(x, ".a.")
```

```{r}
dot <- "\\."
writeLines(dot)
```

```{r}
x <- c("abc", "a.c", "bef")
str_view(x, "a\\.c")
```

```{r}
x <- "a\\b"
writeLines(x)
```

```{r}
str_view(x, "\\\\")
```

### 14.3.1.1 Exercises

1.
* "\\" doesn't match a \\ because \\ is an escape character in strings. 
* "\\\\" doesn't match because it is the equivalent of \\ in the regular expression, which will be an escape character. 
* "\\\\\\" doesn't match because the first two backslashes will match a literal \\, but the third backslash will be an escape character in the string.


### 14.3.2 Anchors

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
```

```{r}
str_view(x, "a$")
```

```{r}
x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")
```

```{r}
str_view(x, "^apple$")
```

### 14.3.2.1 Exercises

1. You would match the literal string "\$^\$" using "\^\\\\\$\\\\^\\\\\$$".


```{r}
y <- c("abc$^$def", "$^$")


str_view(y, "^\\$\\^\\$$")
```

### 14.3.3 Character classes and alternatives

```{r}
x <- c("abc", "a.c", "a*c", "a c")
str_view(x, "a[.]c")
```

```{r}
str_view(x, ".[*]c")
```

```{r}
str_view(x, "a[ ]")
```

```{r}
x <- c("grey", "gray")
str_view(x, "gr(e|a)y")
```

### 14.3.3.1 Exercises

1.
* Start with a vowel:

```{r}
fruits <- c("apple", "banana", "orange", "mango", "apricot", "pear")
str_view(fruits, "^[aeiou]")
```


* Only contain consonants:

```{r}
z <- c("fly", "try", "say", "day", "spy", "pay")
str_view(z, "[aeiou]", match=FALSE)
```

* End with -ed but not with -eed:

```{r}
q <- c("seed", "freed", "bed", "red", "feed", "led")
str_view(q, "(^|[^e])ed$")
```

* End with -ing or -ise

```{r}
p <- c("swing", "exercise", "ring", "early", "wing", "active")
str_view(p, "i(ng|se)$")
```

### 14.3.4 Repetition

```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"

str_view(x, "CC?")
```

```{r}
str_view(x, "CC+")
```

```{r}
str_view(x, 'C[LX]+')
```

```{r}
str_view(x, "C{2}")
```

```{r}
str_view(x, "C{2,}")
```

```{r}
str_view(x, 'C{2,3}?')
```

```{r}
str_view(x, 'C[LX]+?')
```

### 14.3.4.1 Exercises

1.

* '?' is equivalent to {0,1} in {m,n} form. 
* '+' is equivalent to {1,} in {m,n} form. 
* '*' is equivalent to {0,} in {m,n} form.

### 14.3.5 Grouping and backreferences

```{r}
str_view(fruit, "(..)\\1", match=TRUE)
```

### 14.3.5.1 Exercises

1.
* (.)\1\1 - the same character three times in a row

* "(.)(.)\\\\2\\\\1" - a pair of two characters followed by the same two characters in the reverse order

* (..)\1 - any two characters repeated by the same two characters

* "(.).\\\\1.\\\\1" - a character followed by any character followed by the original character followed by any character followed by the original character again

* "(.)(.)(.).*\\\\3\\\\2\\\\1" - three characters followed by zero or more characters of any kind followed by the original three characters in the reverse order

## 14.4 Tools

### 14.4.1 Detect matches

```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")
```

```{r}
sum(str_detect(words, "^t"))
```

```{r}
mean(str_detect(words, "[aeiou]$"))
```

```{r}
no_vowels_1 <- !str_detect(words, "[aeiou]")

no_vowels_2 <- str_detect(words, "^[^aeiou]+$")

identical(no_vowels_1, no_vowels_2)
```

```{r}
words[str_detect(words, "x$")]
```

```{r}
str_subset(words, "x$")
```

```{r}
df <- tibble(
  word = words, 
  i = seq_along(word)
)
df %>% 
  filter(str_detect(word, "x$"))
```

```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
```

```{r}
mean(str_count(words, "[aeiou]"))
```

```{r}
df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )
```

```{r}
str_count("abababa", "aba")
```

```{r}
str_view_all("abababa", "aba")
```

### 14.4.1.1 Exercises

1.

All words that start or end with x:

* Single regular expression:
```{r}
str_subset(words, "^x|x$")
```

* Split into parts:
```{r}
start_with_x <- str_detect(words, "^x")
end_with_x <- str_detect(words, "x$")
words[start_with_x | end_with_x]
```

All words that start with a vowel and end with a consonant:

* Single regular expression:
```{r}
str_subset(words, "^[aeiou].*[^aeiou]") %>% head()
```

* Split into parts:
```{r}
start_with_vowel <- str_detect(words, "^[aeiou]")
end_with_consonant <- str_detect(words, "[^aeiou]$")
words[start_with_vowel & end_with_consonant] %>% head()
```
Any words that contain at least one of each vowel:
```{r}
a <- str_detect(words, "a")
e <- str_detect(words, "e")
i <- str_detect(words, "i")
o <- str_detect(words, "o")
u <- str_detect(words, "u")
words[a & e & i & o & u]
```
There are no words that contain at least one of each vowel.

### 14.4.2 Extract matches

```{r}
length(sentences)
```

```{r}
head(sentences)
```

```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match
```

```{r}
has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)
```

```{r}
more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)
```

```{r}
str_extract(more, colour_match)
```

```{r}
str_extract_all(more, colour_match)
```

```{r}
str_extract_all(more, colour_match, simplify = TRUE)
```

```{r}
x <- c("a", "a b", "a b c")
str_extract_all(x, "[a-z]", simplify = TRUE)
```

### 14.4.2.1 Exercises

1.

```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")
colour_match

more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match, match = TRUE)
```

### 14.4.3 Grouped matches

```{r}
noun <- "(a|the) ([^ ]+)"

has_noun <- sentences %>%
  str_subset(noun) %>%
  head(10)
has_noun %>% 
  str_extract(noun)
```

```{r}
has_noun %>% 
  str_match(noun)
```

```{r}
tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)", 
    remove = FALSE
  )
```

### 14.4.3.1 Exercises

1.

```{r}
number_word <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten) ([^ |\\.]+)"

has_number <- sentences %>%
  str_subset(number_word)

has_number %>% 
  str_extract(number_word)
```

### 14.4.4 Replacing matches

```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
```

```{r}
str_replace_all(x, "[aeiou]", "-")
```

```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```

```{r}
sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```

### 14.4.4.1 Exercises

```{r}
x <- "red/orange/yellow/green/blue/purple"
str_replace_all(x, "/", "\\\\")
```

### 14.4.5 Splitting

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ")
```

```{r}
"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]
```

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)
```

```{r}
fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)
```

```{r}
x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))
```

```{r}
str_split(x, " ")[[1]]
```

```{r}
str_split(x, boundary("word"))[[1]]
```

### 14.4.5.1 Exercises

1.

```{r}
x <- "apples, pears, and bananas"
str_split(x, ", +(and )?")[[1]]
```











